Toward single particle reconstruction without particle picking: Breaking the detection limit

### 2018-08-31 pnas/Towards.tex
Updated Appendix B

Throughout the proof, we consider the case of one signal $K=1$. The extension to $K>1$ is straightforward by averaging the contributions of all signal with  appropriate weights; see~\cite{boumal2017heterogeneous}.

We will let the number of instances of the signal $M$ grow with $N$, and write $M=M_N$ to emphasize this. We assume $M_N$ grows proportionally with $N$, and define: \TODO{---Nico: the paper now no longer references $M_N$ and defines $\gamma$ 'as is' for given $M$ and $N$; may want to adapt things here as a result. It wouldn't make sense to make $M_N$ appear in the paper I think, as it just makes things more complicated. In a sense, much of this beginning of appendix is just repeating things we already said cleanly in the paper: let's just follow up from that.}
%
\begin{equation}
\gamma = \lim_{N\to\infty} \frac{M_NL}{N}<1.
\end{equation}
%
We will assume that $M_N=\Omega(N)$, so that $\gamma>0$. In the sequel, we will suppress the explicit dependence of $M$ on $N$ for notational convenience.

We start by considering the mean of the data:
%
\begin{equation}
a_y^1 = \frac{1}{N}\sum_{i=0}^{N-1} y[i] =
\frac{1}{N/L}\sum_{j=0}^{M-1}\frac{1}{L}\sum_{i=0}^{L-1}x[i] +    
\underbrace{\frac{1}{N}\sum_{i=0}^{N-1}\varepsilon[i]}_{\text{noise term}}
\xrightarrow{a.s.}\gamma a_x^1,
\end{equation}
%
where the noise term converges to zero almost surely (a.s.) by the strong law of large numbers.

We proceed with the (second-order) autocorrelation for fixed $\ell\in[0,\ldots,L-1]$. We can compute:
%
\begin{align*}
%
a_y^2[\ell] & = \frac{1}{N}\sum_{i=0}^{N-1-\ell}y[i]y[i+\ell]
\nonumber \\
& = \underbrace{\frac{1}{N}\sum_{j=1}^{M}\sum_{i=0}^{L-\ell-1}x[i]x[i+\ell]}_{\text{signal term}} + \underbrace{\frac{1}{N}\sum_{i=0}^{N-1-\ell}\varepsilon[i]\varepsilon[i+\ell]}_{\text{noise term}}
\\ & + \underbrace{\frac{\TODO{1,2?}}{N} \sum_{j=1}^{M} \sum_{i=0}^{L-1} x[i] \varepsilon[s_j + i + \ell]}_{\text{cross-term}}. 
%
\end{align*}

The cross-terms are linear in the noise, and are easily shown to vanish almost surely in the limit $N\to\infty$, by the strong law of large numbers. As for the signal term, we break it into $M$ different sums, each containing one copy of the signal. This gives:
%
\begin{equation} \label{eq:2nd_moment_signal_term}
%
\frac{1}{N}\sum_{j=1}^{M}\sum_{i=0}^{L-\ell-1}x[i]x[i+\ell] = \frac{ML}{N}\frac{1}{L}\sum_{i=0}^{L-\ell-1}x[i]x[i+\ell]\xrightarrow{N\to\infty}\gamma a_x^2[\ell].
%
\end{equation}
%

We next analyze the pure noise term. When $\ell\neq 0$, we can break the noise term into a sum of independent terms:
%
\begin{equation}
%
\frac{1}{N}\sum_{i=0}^{N-1-\ell} \varepsilon[i]\varepsilon[i+\ell] = \frac{1}{\ell}\sum_{i=0}^{\ell-1}\frac{1}{N/\ell}\sum_{j=0}^{N/\ell -1} \varepsilon[j\ell + i] \varepsilon[(j+1)\ell + i].
%
\end{equation}
%
Each sum $\frac{1}{N/\ell}\sum_{j=0}^{N/\ell -1} \varepsilon[j\ell + i] \varepsilon[(j+1)\ell + i]$ is an average of $N/\ell$ independent terms with expectation zero, and thus converges to zero almost surely as $N\to\infty$. If $\ell=0$, then we have:
%
\begin{equation}
%
\frac{1}{N}\sum_{i=0}^{N-1} \varepsilon^2[i] \xrightarrow{a.s.} \sigma^2.
%
\end{equation}

>>>>

%Throughout the proof, we consider the case of one signal $K=1$. The extension to $K>1$ is straightforward by averaging the contributions of all signal with  appropriate weights; see~\cite{boumal2017heterogeneous}. 
%We will let the number of instances of the signal $M$ grow with $N$, and write $M=M_N$ to emphasize this.
We consider the asymptotic regime $N,M\to\infty$ and assume that $M=\Omega(N)$, so that
% We assume $M_N$ grows proportionally with $N$, and recall that
% define: %\TODO{---Nico: the paper now no longer references $M_N$ and defines $\gamma$ 'as is' for given $M$ and $N$; may want to adapt things here as a result. It wouldn't make sense to make $M_N$ appear in the paper I think, as it just makes things more complicated. In a sense, much of this beginning of appendix is just repeating things we already said cleanly in the paper: let's just follow up from that.}
%
\begin{equation*}
\gamma := \frac{ML}{N}>0.
\end{equation*}
%
%We will consider the regime $N,M\to\infty$ and assume that $M_N=\Omega(N)$, so that $\gamma>0$. %In the sequel, we will suppress the explicit dependence of $M$ on $N$ for notational convenience.
We start by considering the mean of the data:
%
\begin{equation}
a_y^1 = \frac{1}{N}\sum_{i=0}^{N-1} y[i] =
\frac{1}{N/L}\sum_{j=0}^{M-1}\frac{1}{L}\sum_{i=0}^{L-1}x[i] +    
\underbrace{\frac{1}{N}\sum_{i=0}^{N-1}\varepsilon[i]}_{\text{noise term}}
\xrightarrow{a.s.}\gamma a_x^1,
\end{equation}
%
where the noise term converges to zero almost surely (a.s.) by the strong law of large numbers.

We proceed with the (second-order) autocorrelation for fixed $\ell\in[0,\ldots,L-1]$. We can compute:
%
\begin{align*}
%
a_y^2[\ell] & = \frac{1}{N}\sum_{i=0}^{N-1-\ell}y[i]y[i+\ell]
\nonumber \\
& = \underbrace{\frac{1}{N}\sum_{j=1}^{M}\sum_{i=0}^{L-\ell-1}x[i]x[i+\ell]}_{\text{signal term}} + \underbrace{\frac{1}{N}\sum_{i=0}^{N-1-\ell}\varepsilon[i]\varepsilon[i+\ell]}_{\text{noise term}}
\\ & + \underbrace{\frac{2}{N} \sum_{j=1}^{M} \sum_{i=0}^{L-1} x[i] \varepsilon[s_j + i + \ell]}_{\text{cross-term}}. 
%
\end{align*}

The cross-terms are linear in the noise, and are easily shown to vanish almost surely in the limit $N\to\infty$, by the strong law of large numbers. As for the signal term, we break it into $M$ different sums, each containing one copy of the signal. This gives:
%
\begin{align} \label{eq:2nd_moment_signal_term}
%
\frac{1}{N}\sum_{j=1}^{M}\sum_{i=0}^{L-\ell-1}x[i]x[i+\ell] & = \frac{ML}{N}\frac{1}{L}\sum_{i=0}^{L-\ell-1}x[i]x[i+\ell]\nonumber\\&\xrightarrow{N\to\infty}\gamma a_x^2[\ell].
%
\end{align}
%

We next analyze the pure noise term. When $\ell\neq 0$, we can break the noise term into a sum of independent terms:
%
\begin{equation}
%
\frac{1}{N}\sum_{i=0}^{N-1-\ell} \varepsilon[i]\varepsilon[i+\ell] = \frac{1}{\ell}\sum_{i=0}^{\ell-1}\frac{1}{N/\ell}\sum_{j=0}^{N/\ell -1} \varepsilon[j\ell + i] \varepsilon[(j+1)\ell + i].
%
\end{equation}
%
Each sum $\frac{1}{N/\ell}\sum_{j=0}^{N/\ell -1} \varepsilon[j\ell + i] \varepsilon[(j+1)\ell + i]$ is an average of $N/\ell$ independent terms with expectation zero, and thus converges to zero almost surely as $N\to\infty$. If $\ell=0$, then we have:
%
\begin{equation*}
%
\frac{1}{N}\sum_{i=0}^{N-1} \varepsilon^2[i] \xrightarrow{a.s.} \sigma^2.
%
\end{equation*}
